\mpx{}: mixed precision Training for JAX

### 2025-05-29 doc/paper/main.tex
continued doc

In recent years, JAX~\cite{deepmind2020jax} has become an increasingly popular tool for training neural networks. 
A key feature of JAX is its just-in-time compilation, which significantly accelerates code execution on hardware accelerators such as GPUs and TPUs. 
In addition, JAX supports sharding mechanisms for multi-device and multi-node training, making it easy to scale up the training of large models across entire computing clusters.

Many libraries have been developed to extend JAX's capabilities, including Equinox, Flax, and Optax~\citep{kidger2021equinox,flax2020github,deepmind2020jax}. 
However, JAX has so far lacked a reliable, up-to-date library to enable mixed-precision training. The only existing option, JMP\citep{jmp}, is no longer maintained and cannot handle arbitrary JAX PyTrees, which are commonly used in modern neural-network toolboxes like Equinox~\citep{kidger2021equinox}.

Mixed precision training is indispensable for efficient training of large models~\cite{mixed_precision_paper}. 
By executing both the forward and backward passes of a neural network in half-precision floating-point format (requiring only \qty{16}{\bit}), mixed-precision techniques reduce memory requirements, particularly for intermediate activations, and speed up training. 
This speedup arises from lower memory transfer overhead and from specialized half-precision tensor cores in case the used hardware features them. 
Crucially, these benefits typically come without compromising accuracy.

This work presents \mpx{}, a mixed-precision training toolbox for JAX. 
\mpx{} builds on the JMP implementation and leverages Equinox's flexibility to overcome limitations in handling arbitrary PyTrees. It further introduces gradient transformations and optimizer wrappers, enabling seamless conversion of a full-precision pipeline into mixed precision with only minor modifications. \mpx{}'s source code, documentation, and usage examples are available at \url{https://github.com/AlexGraefe/mixed_precision_for_JAX}.

Before detailing the core functionality of \mpx{}, we first provide an overview of mixed-precision training. 
We then illustrate how \mpx{} can be integrated into existing workflows by presenting sample code that demonstrates its main features.

\section{Basics of mixed precision Training}
This section summarizes the original mixed precision method from NVIDIA's Automatic mixed precision \citep{mixed_precision_paper} and the paper by Micikevicius et al. \citep{mixed_precision_paper}.

mixed precision training involves performing most computations in the forward and backward passes of a neural network using 16-bit floating-point numbers. This approach offers two main advantages:
\begin{itemize}
    \item Reduced GPU memory usage (approximately 50\% compared to full precision)
    \item Potential speedup through decreased memory access times and utilization of specialized half-precision tensor cores of TPUs or GPUs.
\end{itemize}

\subsection{Loss Scaling}
A critical component of successful mixed precision training is loss scaling. Due to float16's decreased resolution, small gradients may be cast to zero, negatively impacting training performance. The solution involves:
\begin{enumerate}
    \item Scaling the loss by a factor $> 1$
    \item Calculating gradients with the scaled loss
    \item Converting gradients to float32
    \item Dividing by the scaling factor to obtain original gradients
\end{enumerate}

The scaling factor can be automatically adjusted using a simple heuristic:
\begin{itemize}
    \item If scaled gradients exceed float16 range (inf), reduce scaling and skip model update.
    \item If scaled gradients remain within float16 range for an extended period, increase scaling.
\end{itemize}

@@

The \mpx{} library provides essential transformations for mixed precision training while retaining JAX's low-level flexibility. 
Its design leverages JAX's type-promotion behavior\footnote{\url{https://docs.jax.dev/en/latest/jep/9407-type-promotion.html}}, ensuring that operations are performed in the precision to which the inputs and outputs have been cast—provided that all constants in the function are positioned on the left side of the type-promotion lattice. 
As such, \mpx{} primarily focuses on casting function inputs and outputs to the desired precision. 
Concretely, it first defines functions to cast PyTrees (Section~\ref{sec:castingpytrees}), then uses these to cast both functions (Section~\ref{sec:castingfunctions}) and gradient calculations (Section~\ref{sec:castinggradients}) to mixed precision. In addition, we employ dynamic loss scaling~\citep{mixed_precision_paper} (Section~\ref{sec:scaling}) to stabilize mixed precision training.

@@

\mpx{} provides several functions for casting arbitrary PyTrees to a desired floating-point data type. These include
\texttt{cast\_tree(tree, dtype)}, \texttt{cast\_to\_half\_precision(x)}, \texttt{cast\_to\_float16(x)}, \texttt{cast\_to\_bfloat16(x)}, and \texttt{cast\_to\_float32(x)}. All of these functions inspect each leaf of the input PyTree, and any leaf that is a JAX array of a floating type is cast to the specified data type. Other leaves, such as integer arrays, remain unchanged.
Excluding integer arrays is important to avoid accidental casting of random generator keys.

@@

\begin{itemize}
    \item \texttt{scale(x)}: Multiplies all floating-point leaves of the PyTree \texttt{x} by the current scaling factor.
    \item \texttt{unscale(x)}: Divides all floating-point leaves of \texttt{x} by the scaling factor and casts them to full precision.
    \item \texttt{adjust(grads\_finite)}: Updates the scaling factor based on the stability of the gradients, following the techniques described by \cite{nvidia_amp}.
\end{itemize}

Because \texttt{DynamicLossScaling} inherits from \texttt{eqx.Module}, it is itself a PyTree. 
Consequently, it can be used within \texttt{jit}-compiled functions and also replicated or sharded across multiple devices for distributed training.

@@

\mpx{} provides Equinox-compatible versions of gradient-calculation functions, namely \texttt{filter\_grad(func, scaling, has\_aux=False, use\_mixed\_precision=True)} and \texttt{filter\_value\_and\_grad(func, scaling, has\_aux=False, use\_mixed\_precision=True)}. 
These functions compute gradients with mixed precision and loss scaling (as described in Section~\ref{sec:scaling}) and also automatically adjust the loss scaling value. 
They serve as drop-in replacements for the corresponding Equinox functions, allowing users to reuse existing training pipelines with minimal changes (cf. Section~\ref{sec:eqxvsmpxmodelupdate}).

@@

\begin{enumerate}
    \item Cast all input arguments to half precision (\texttt{float16} or \texttt{bfloat16}).
    \item Call the original function (model forward pass and loss).
    \item Scale the function's output by the current dynamic scaling factor.
    \item Apply \texttt{equinox.filter\_grad} to obtain scaled gradients.
    \item Unscale the gradients by dividing them by the scaling factor.
    \item Check whether the gradients are finite.
    \item Update the loss scaling based on the gradient's finiteness.
    \item Return a tuple containing:
    \begin{itemize}
        \item The updated \texttt{scaling} object.
        \item A boolean indicating if the gradients are finite (for the optimizer step).
        \item The computed gradients, whose format is a PyTree with the same format as the model parameters.
        \item Auxiliary values (if \texttt{has\_aux=True}).
    \end{itemize}
\end{enumerate}

\subsection{Optimizer}

Because MPX produces gradients as PyTrees, they are directly compatible with Optax optimizers~\cite{deepmind2020jax}. 
However, as discussed above, it may be necessary to skip optimizer updates altogether if the gradients become infinite due to loss scaling. 
The function \texttt{optimizer\_update(model, optimizer, optimizer\_state, grads, grads\_finite)} implements this logic and updates the model only when gradients are finite.
Consequently, instead of calling \texttt{optimizer.update} followed by \texttt{eqx.apply\_updates} (as in a typical Equinox training pipeline), one can simply call \texttt{mpx.optimizer\_update}.

\section{Example}
Here, we provide an example and show which parts in a training pipeline need to be changed for mixed precision training.

@@

For the largest part, the implementation of the model must not be changed. 
As \mpx{} works with arbitrary PyTrees, every Toolbox that defines their model/parameters as PyTrees, like Flax~\citep{flax2020github} or Equinox~\citep{kidger2021equinox} can be used in conjunction with \mpx{} (Listing~\ref{ex:model}). 
Only parts of the model that contain critical operations with a high risk of overflows, like sum or mean, must be transformed to full precision via \texttt{mpx.force\_full\_precision}.

\subsection{Model Updates}
\label{sec:eqxvsmpxmodelupdate}
Model updates via \mpx{} are very similar to model updates via Equinox (Figure~\ref{ex:model}). Gradients can be calulated via

@@

\section{Evaluation}
We evaluate \mpx{} on two different machines: A mid-level desktop PC (AMD Ryzen 9 5950X, \qty{82}{\giga\byte} RAM, Nvidia GeForce RTX4070 GPU with \qty{12}{\giga\byte} VRAM) and a node on a high-performance cluster (CLAIX-2023, Intel Xeon 8468 Sapphire CPUs, \qty{512}{\giga\byte} RAM, four NVIDIA H100 GPUs with \qty{96}{\giga\byte} VRAM each connected via NVLink).
On paper, the desktop PC's GPU half precision computation speed is the same for half and full precision, while the H100 GPUs has double the speed for half than for full precision.

We train small vision transformers (ViT)~\cite{dosovitskiy2020image,steiner2021train}.
On both machines, the transformer has a feature size is 256, the residual blocks have one hidden layer with 800 neurons.
The desktop PC trains the ViT on Cifar100 and the high-performance cluster on ImageNet1k, equally dividing each batch between the GPUs using a data-parallel approach.
We train without gradient accumulation, which would be another way to optimize runtime and memory consumption.
We vary the number of batches and measure the time the machines need to for each training step on the entire dataset, excluding the times used for data loading.

We measure the RAM the program uses by turning off automatic preallocation and setting ``XLA\_PYTHON\_CLIENT\_PREALLOCATE'' to ``platform'', which minimizes the VRAM consumption of JAX\footnote{\url{https://docs.jax.dev/en/latest/gpu_memory_allocation.html}}
For training speed, we turn both off.

@@

        % Second graph: Mixed Precision
        \addplot[
            color=red,
            mark=square*,
            thick,
            dashed
        ] coordinates {
            (128, 12.6)
            (256, 12.0)
            (512, 11.4)
            (768, 11.6)
            (1024, 11.5)
        };
        \addlegendentry{Mixed Precision}
        \end{axis}
    \end{tikzpicture}
    \caption{Desktop Computer (RTX4070)}
    \end{subfigure}
    \begin{subfigure}[t]{0.49\linewidth}
    \begin{tikzpicture}
        \begin{axis}[
            width=0.8\textwidth,
            height=7cm,
            xlabel={Batch Size},
            ylabel={Training Time per Epoch (s)},
            legend pos=south east,
            grid=major,
            grid style={dashed,gray!30},
            xmin=0,
            ymin=0,
        ]
        % First graph: Full Precision
        \addplot[
            color=blue,
            mark=o,
            thick
        ] coordinates {
            (128, 19.7)
            (256, 18.8)
            (512, 19.7)
        };
        \addlegendentry{Full Precision}

        % Second graph: Mixed Precision
        \addplot[
            color=red,
            mark=square*,
            thick,
            dashed
        ] coordinates {
            (128, 12.6)
            (256, 12.0)
            (512, 11.4)
            (768, 11.6)
            (1024, 11.5)
        };
        \addlegendentry{Mixed Precision}
        \end{axis}
    \end{tikzpicture}
    \caption{High-Performance Cluster ($4\times$ H100)}
    \end{subfigure}
    \caption{Comparison of training step times for full precision and mixed precision as a function of the number of batches.}
    \label{fig:timing}
\end{figure}

Mixed precision training via \mpx{} reduces both RAM and processing times.
For the desktop PC, the RAM is reduced by factor $1.8\times$ and processing time by $1.7\times$.
The speedup is caused by reduced memory loading times as the RTX4070 features no computing speedup for half precision.

>>>>

In recent years, JAX~\citep{deepmind2020jax} has emerged as a popular tool for training neural networks. 
A standout feature of JAX is its just-in-time compilation, which significantly accelerates code execution on hardware accelerators such as GPUs and TPUs. 
Additionally, JAX supports scalable multi-device and multi-node training across computing clusters through user-friendly and efficient sharding mechanisms.

To extend JAX's capabilities, several libraries have been developed, including Equinox, Flax, and Optax~\citep{kidger2021equinox,flax2020github,deepmind2020jax}. 
Despite these advancements, one critical area where JAX falls short is in supporting mixed-precision training, a method essential for efficiently handling large models~\cite{mixed_precision_paper}. 
Although JMP~\citep{jmp} was previously available for this purpose, it is no longer maintained and cannot process arbitrary JAX PyTrees commonly used in modern neural-network toolboxes.

Mixed-precision training plays a pivotal role in optimizing the performance of large model training by executing both forward and backward passes in half-precision floating-point format (\qty{16}{\bit}). 
This technique significantly reduces memory requirements for intermediate activations and speeds up training by minimizing memory transfer overheads while utilizing specialized half-precision tensor cores (when supported by hardware), all without compromising accuracy~\cite{mixed_precision_paper}.

To facilitate the straightforward implementation of mixed-precision training in JAX, we introduce \mpx{}, a mixed-precision training toolbox tailored for JAX. 
\mpx{} builds upon the JMP implementation and leverages Equinox's flexibility to overcome limitations associated with handling arbitrary PyTrees. 
It further introduces gradient transformations and optimizer wrappers that enable seamless conversion from a full-precision pipeline to mixed precision with minimal modifications. The source code, documentation, and usage examples for \mpx{} are accessible at \url{https://github.com/AlexGraefe/mixed_precision_for_JAX}.

Before explaining the core functionality of \mpx{}, we provide an overview of mixed-precision training principles. 
Subsequently, we demonstrate how \mpx{} can be integrated into existing workflows through sample code illustrating its main features.
Finally, we evaluate the performance of \mpx{} in the context of training vision transformers.

\section{Basics of mixed precision Training}
In this section, we summarize the foundational principles of mixed precision training as outlined in the original method~\citep{mixed_precision_paper}. 
Mixed precision training involves executing most computations during the forward and backward passes of a neural network using 16-bit floating-point numbers. 
This approach offers two primary advantages: 
\begin{itemize} 
    \item Reduced GPU Memory Usage: Mixed precision can decrease memory usage by approximately \SI{50}{\percent} compared to full precision. 
    \item Potential Speedup: This technique can lead to faster execution due to reduced memory access times and the utilization of specialized half-precision tensor cores available in TPUs or GPUs. 
\end{itemize}

\subsection{Loss Scaling}
One downside of using half-precision floating-point numbers is their decreased resolution. 
Consequently, small gradients may be rounded to zero, which can negatively impact training performance.

To mitigate these effects, loss scaling has been proposed. 
The main idea involves multiplying the loss by a factor greater than one and calculating the gradient of this scaled loss. 
Consequently, the gradients are also scaled by this factor, making them larger and allowing them to be represented more accurately by half-precision numbers.

However, selecting a constant scaling factor poses challenges. 
A small scaling factor might still result in gradients being rounded to zero, while an excessively large scaling factor can cause overflows where gradients become infinite. 
Additionally, different stages of training may benefit from varying scaling factors depending on the training state.

Dynamic loss scaling addresses this issue by adapting the scaling factor at runtime.
Its objective is to maximize the scaling factor for best accuracy with half-precision numbers while avoiding overflow. 
This is achieved through a simple heuristic that periodically increases the scaling factor and decreases it when gradients turn infinite.

In summary, mixed precision training with dynamic loss scaling involves the following steps:
\begin{enumerate}
    \item Calculate the prediction of the model in half-precision.
    \item Calculate the loss and scale it by a factor greater than 1.
    \item Calculate gradients of the scaled loss with respect to the model parameters.
    \item Converting gradients to float32.
    \item Divide them by the scaling factor to retrieve original gradients.
    \item Update the scaling factor:
    \begin{enumerate}[label=(\alph*)]
        \item If scaled gradients exceed float16's representable range (resulting in infinity), reduce the scaling and skip updating model parameters. 
        \item If scaled gradients remain within float16 range for an extended period, increase scaling.
    \end{enumerate}
    \item Update the model parameters.
\end{enumerate}

The primary goal of \mpx{} is to offer functionalities that streamline the implementation of all these steps in mixed precision training. 
In the following section, we provide an overview of these features.

@@

The \mpx{} library offers essential transformations for mixed precision training while preserving JAX's low-level flexibility. 
Similar to JMP~\cite{jmp}, implementationts design leverages JAX's type-promotion behavior\footnote{\url{https://docs.jax.dev/en/latest/jep/9407-type-promotion.html}}, ensuring that operations are executed in the precision to which inputs and outputs have been cast—provided all constants in the function are positioned on the left side of the type-promotion lattice. 
Consequently, \mpx{} primarily focuses on casting function inputs and outputs to the desired precision.

Specifically, \mpx{} first defines functions for casting PyTrees (Section\ref{sec:castingpytrees}), then utilizes these functions to cast both general functions (Section\ref{sec:castingfunctions}) and gradient calculations (Section\ref{sec:castinggradients}) into mixed precision. 
The gradient calculations include dynamic loss scaling (Section~\ref{sec:scaling}) to stabilize mixed precision training.
Moreover, \mpx{} provides wrappers around optimizers to exclude update steps where gradients are infinite (Section~\ref{sec:optimizer}).

@@

\mpx{} provides several functions for casting arbitrary PyTrees to a desired floating-point data type. These functions include:
\texttt{cast\_tree(tree, dtype)}, \texttt{cast\_to\_half\_precision(x)}, \texttt{cast\_to\_float16(x)}, \texttt{cast\_to\_bfloat16(x)}, and \texttt{cast\_to\_float32(x)}. 
Each function inspects the leaves of the input PyTree, casting any leaf that is a JAX array with a floating-point type to the desired data type. 
Other leaves, such as integer arrays, remain unchanged. 
Excluding integer arrays is crucial to prevent accidental casting of random generator keys.

@@

\begin{itemize}
    \item \texttt{scale(x)}: Multiplies all floating-point leaves of the PyTree \texttt{x} by the current scaling factor.
    \item \texttt{unscale(x)}: Divides all floating-point leaves of \texttt{x} by the scaling factor and casts them to full precision.
    \item \texttt{adjust(grads\_finite)}: Updates the scaling factor based on the stability of the gradients, following the techniques described by~\cite{mixed_precision_paper}.
\end{itemize}

Because \texttt{DynamicLossScaling} inherits from \texttt{eqx.Module}, it is itself a PyTree. 
Consequently, it can be used within \texttt{jit}-compiled functions and also be used for multi-device and multi-node training using replicated sharding.

@@

\mpx{} provides Equinox-compatible versions of gradient-calculation functions: \texttt{filter\_grad(func, scaling, has\_aux=False, use\_mixed\_precision=True)} and \texttt{filter\_value\_and\_grad(func, scaling, has\_aux=False, use\_mixed\_precision=True)}. 
These functions compute gradients using mixed precision and loss scaling (as described in Section\ref{sec:scaling}) while automatically adjusting the loss scaling value. 
They act as drop-in replacements for the corresponding Equinox functions, enabling users to reuse existing training pipelines with minimal modifications (cf. Section\ref{sec:eqxvsmpxmodelupdate}).

@@

\begin{enumerate}
    \item Cast all input arguments to half precision (\texttt{float16} or \texttt{bfloat16}).
    \item Execute the original function (model forward pass and loss).
    \item Scale the function's output by the current dynamic scaling factor.
    \item Apply \texttt{equinox.filter\_grad} to obtain scaled gradients.
    \item Unscale the gradients by dividing them by the scaling factor.
    \item Check whether the gradients are finite.
    \item Update the loss scaling based on the finiteness of gradients.
    \item Return a tuple containing:
    \begin{itemize}
        \item The updated \texttt{scaling} object.
        \item A boolean indicating if the gradients are finite (for the optimizer step).
        \item The computed gradients, whose format is a PyTree with the same format as the model parameters.
        \item Auxiliary values (if \texttt{has\_aux=True}).
    \end{itemize}
\end{enumerate}

\subsection{Optimizer}
\label{sec:optimizer}

Since \mpx{} produces gradients as PyTrees, they are directly compatible with Optax optimizers~\cite{deepmind2020jax}.
However, as previously discussed, it may be necessary to skip optimizer updates if the gradients become infinite due to loss scaling. 
The function \texttt{optimizer\_update(model, optimizer, optimizer\_state, grads, grads\_finite)} implements this logic by updating the model only when gradients are finite.

Thus, instead of calling \texttt{optimizer.update} followed by \texttt{eqx.apply\_updates}, as in a typical Equinox training pipeline, one can simply use \texttt{mpx.optimizer\_update}.

\section{Example}
In this section, we provide an example to demonstrate the changes required in a training pipeline for mixed precision training via \mpx{}.

@@

@@

For the largest part, the implementation of the model must not be changed.
As \mpx{} works with arbitrary PyTrees, every Toolbox that defines their model/parameters as PyTrees, like Flax~\citep{flax2020github} or Equinox~\citep{kidger2021equinox} can be used in conjunction with \mpx{} (Example~\ref{ex:model}). 
Only parts of the model that contain critical operations with a high risk of overflows, like sum or mean, must be transformed to full precision via \texttt{mpx.force\_full\_precision}.

\subsection{Model Updates}
\label{sec:eqxvsmpxmodelupdate}
Model updates via \mpx{} closely resemble those performed with Equinox (Example~\ref{ex:update}). 
Instead of calling \texttt{eqx.filter\_grad}, \texttt{mpx.filter\_grad} is used. Similarly, \texttt{mpx.optimizer\_update} replaces invoking the Optax optimizer followed by a model update.
The remainder of the training step—such as loading batches, sharding the model and batch—remains unchanged. 
All these steps can be compiled using tools like \texttt{eqx.filter\_jit}.

\section{Evaluation}
We evaluate \mpx{} on two distinct machines: a mid-level desktop PC equipped with an AMD Ryzen 9 5950X, \qty{82}{\giga\byte} RAM, and an Nvidia GeForce RTX4070 GPU featuring \qty{12}{\giga\byte} VRAM; and a node from a high-performance cluster (CLAIX-2023) comprising Intel Xeon 8468 Sapphire CPUs, \qty{512}{\giga\byte} RAM, and four NVIDIA H100 GPUs connected via NVLink each with \qty{96}{\giga\byte} VRAM. 
On paper, the desktop PC's GPU maintains equivalent computation speeds for both half and full precision, whereas the H100 GPUs deliver double the speed for half precision compared to full precision.

We train vision transformers (ViT)~\cite{dosovitskiy2020image,steiner2021train}. On the desktop PC, the transformer features a size of 256 with residual blocks containing one hidden layer of 800 neurons. T
The cluster's transformer mirrors ViT-Base dimensions, with a feature size of 768 and residual blocks having one hidden layer of 3072 neurons.

The desktop PC trains the ViT on CIFAR100 while the high-performance cluster handles ImageNet1k, dividing each batch equally across GPUs using a data-parallel approach. We train without gradient accumulation, which could further optimize runtime and memory consumption

We vary the number of batches and measure the time required for each training step across the entire dataset, excluding data loading times. 
RAM usage is measured by disabling automatic preallocation and setting "XLA\_PYTHON\_CLIENT\_PREALLOCATE" to "platform," minimizing JAX's VRAM consumption\footnote{\url{https://docs.jax.dev/en/latest/gpu_memory_allocation.html}}. 
For training speed assessment, both settings are turned off.

@@

        % Second graph: Mixed Precision
        \addplot[
            color=red,
            mark=square*,
            thick,
            dashed
        ] coordinates {
            (128, 12.6)
            (256, 12.0)
            (512, 11.4)
            (768, 11.6)
            (1024, 11.5)
        };
        \addlegendentry{Mixed Precision}
        \end{axis}
    \end{tikzpicture}
    \caption{Desktop Computer (RTX4070)}
    \end{subfigure}
    \begin{subfigure}[t]{0.49\linewidth}
    \begin{tikzpicture}
        \begin{axis}[
            width=0.8\textwidth,
            height=7cm,
            xlabel={Batch Size},
            ylabel={Training Time per Epoch (s)},
            legend pos=south east,
            grid=major,
            grid style={dashed,gray!30},
            xmin=0,
            ymin=0,
        ]
        % First graph: Full Precision
        \addplot[
            color=blue,
            mark=o,
            thick
        ] coordinates {
            (512, 595.5616071701049)
            (1024, 452.3)
            (2048, 428.24)
            % (4096, 153.62)
            % (5120, 152)
            % (6144, 149.53)
            % (7168, 153.88)
        };
        \addlegendentry{Full Precision}

        % Second graph: Mixed Precision
        \addplot[
            color=red,
            mark=square*,
            thick,
            dashed
        ] coordinates {
            (512, 435.09)
            (1024, 325.61)
            (2048, 272.48)
            % (4096, 117.6)
            % (5120, 113.62)
            % (6144, 109.06)
            % (7168, 104.51)
            % (8192, 102.68)
        };
        \addlegendentry{Mixed Precision}
        \end{axis}
    \end{tikzpicture}
    \caption{High-Performance Cluster ($4\times$ H100)}
    \end{subfigure}
    \caption{Comparison of training step times for full precision and mixed precision as a function of the number of batches.}
    \label{fig:timing}
\end{figure}

Mixed precision training via \mpx{} reduces both RAM and processing times.
For the desktop PC, the RAM is reduced by factor $1.8\times$ and processing time by $1.7\times$.
The speedup is caused by reduced memory loading times as the RTX4070 streaming-multiprocessors feature no computing speedup for half precision.
The cluster features a similar behavior, where both RAM and processing times are reduced.
RAM is reduced by factor $1.8\times$ and processing time by up to $1.57\times$.

These experiments demonstrate that mixed precision via \mpx{} allows to significantly speed up training.
Moreover, it significantly reduces VRAM usage.
